---
title: "Muller corals Alpha Diversity Group Significance"
author: "Becca Maher"
date: "12/10/2019"
output: html_document
---
## Script for calculating alpha diversity statistics and testing for significant differences with metadata variables

### 1 Setup the working space
```{r libraries, include=FALSE}
rm(list=ls())
library(phyloseq)
library(multcomp)
library(picante)
library(ggplot2 )
library(tidyverse)
library(FSA)
library(MASS)
library(lme4)
library(lmerTest)
library(emmeans)
library(MuMIn)
library(nlme)
```

#### Functions
```{r functions}
# Arc-sine transformation
asinTransform <- function(p) { asin(sqrt(p)) }
# Calculate standard error
sderr <- function(x) {sd(x)/sqrt(length(x))}
#+++++++++++++++++++++++++
# Function to calculate the mean and the standard deviation
# for each group
#+++++++++++++++++++++++++
# data : a data frame
# varname : the name of a column containing the variable
#to be summariezed
# groupnames : vector of column names to be used as
# grouping variables
data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sderr(x[[col]]), na.rm=TRUE)
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
  return(data_sum)
}

normality.plots <- function(x) {
  par(mfrow=c(2,2))
  hist(residuals(x), main = "Histogram", xlab = "Values")
  boxplot(residuals(x), main = "Boxplot", ylab = "Values")
  qqnorm(residuals(x))
  qqline(residuals(x))
  plot(density(residuals(x)), main = "Kernel Density Estimate")
}
```

#### Load the data
Load the alphadiv.csv files created in the RAPID_alpha_statsA.Rmd script. First starting with all species
```{r load}
alphadiv <- read.csv("../data/alphadiv8692.csv")
alphadiv <- read.csv("../data/alphadiv_norick_rar.csv")
head(alphadiv)
```


### 2 Data Summary statistics

```{r}
# Data summaries
# Observed
data_summary(alphadiv, varname = "observed", groupnames = c("bleach"))
data_summary(alphadiv, varname = "observed", groupnames = c("geno"))
data_summary(alphadiv, varname = "observed", groupnames = c("bleach.geno"))
# Simpson
data_summary(alphadiv, varname = "simpson", groupnames = c("bleach"))
data_summary(alphadiv, varname = "simpson", groupnames = c("geno"))
data_summary(alphadiv, varname = "simpson", groupnames = c("bleach.geno"))
# Shannon
data_summary(alphadiv, varname = "shannon", groupnames = c("bleach"))
data_summary(alphadiv, varname = "shannon", groupnames = c("geno"))
data_summary(alphadiv, varname = "shannon", groupnames = c("bleach.geno"))
# FaithPD
data_summary(alphadiv, varname = "faithPD", groupnames = c("bleach"))
data_summary(alphadiv, varname = "faithPD", groupnames = c("geno"))
data_summary(alphadiv, varname = "faithPD", groupnames = c("bleach.geno"))
```
### 3 Non Parametric Tests for significance
Using Kruskal Wallis tests

#### Observed ASVs
```{r}
#Kruskal Wallis tests
kruskal.test(observed ~ bleach, data = alphadiv)
kruskal.test(observed ~ geno, data = alphadiv)
kruskal.test(observed ~ type, data = alphadiv)
#kruskal.test(observed ~ bleach.geno, data = alphadiv)
#pairwise.wilcox.test(alphadiv$observed, alphadiv$bleach.geno, p.adjust.method = 'fdr')
```

#### Simpson's Index
```{r}
kruskal.test(simpson ~ bleach, data = alphadiv)
kruskal.test(simpson ~ geno, data = alphadiv)
kruskal.test(simpson ~ type, data = alphadiv)
#pairwise.wilcox.test(alphadiv$simpson, alphadiv$bleach.geno, p.adjust.method = 'fdr')
```

#### Shannon's Index
```{r}
kruskal.test(shannon ~ bleach, data = alphadiv)
kruskal.test(shannon ~ geno, data = alphadiv)
kruskal.test(shannon ~ type, data = alphadiv)
#pairwise.wilcox.test(alphadiv$shannon, alphadiv$bleach.geno, p.adjust.method = 'fdr')
```

#### Faith PD
```{r}
kruskal.test(faithPD ~ bleach, data = alphadiv)
kruskal.test(faithPD ~ geno, data = alphadiv)
kruskal.test(faithPD ~ type, data = alphadiv)
#pairwise.wilcox.test(alphadiv$faithPD, alphadiv$bleach.geno, p.adjust.method = 'fdr')
```
### Tests for differences within resistant genotypes
```{r}
alphadiv_res <- subset(alphadiv, alphadiv$type == "resistant")
kruskal.test(observed ~ bleach, data = alphadiv_res)
kruskal.test(simpson ~ bleach, data = alphadiv_res)
kruskal.test(shannon ~ bleach, data = alphadiv_res)
kruskal.test(faithPD ~ bleach, data = alphadiv_res)
```

### 4 Parametric tests for significance
Looking for normality and homogeneity of group dispersions
```{r}
# observed ASVs
histogram(alphadiv$observed)
qqnorm(alphadiv$observed)
shapiro.test(alphadiv$observed)
# sqrt transform makes the data normal
histogram(sqrt(alphadiv$observed))
qqnorm(sqrt(alphadiv$observed))
shapiro.test(sqrt(alphadiv$observed))
# test for homogeneity of group dispersions
bartlett.test(sqrt(observed) ~ bleach, data = alphadiv)
bartlett.test(sqrt(observed) ~ geno, data = alphadiv)
## not significant indicates that the variances in each of the groups are the same

# proceeding with anova tests
summary(aov(sqrt(observed) ~ geno*bleach, alphadiv)) # results disagree with 
summary(aov(sqrt(observed) ~ type, alphadiv))
#the kruskal wallis test
TukeyHSD(aov(sqrt(observed) ~ geno, alphadiv)) # this gets wonky because too many comparisons
```

```{r}
# Shannon
histogram(alphadiv$shannon)
qqnorm(alphadiv$shannon)
shapiro.test(alphadiv$shannon)
# sqrt transform makes the data normal
histogram(log(alphadiv$shannon))
qqnorm(log(alphadiv$shannon))
shapiro.test(log(alphadiv$shannon))
## can't seem to find a transformation that makes this normal
## going to stick with kruskal wallis
```