---
title: "Muller corals Alpha Diversity Group Significance"
author: "Becca Maher"
date: "12/10/2019"
output: html_document
---
## Script for calculating alpha diversity statistics and testing for significant differences with metadata variables

### 1 Setup the working space
```{r libraries, include=FALSE}
rm(list=ls())
library(phyloseq)
library(multcomp)
library(picante)
library(ggplot2 )
library(tidyverse)
library(FSA)
library(MASS)
library(lme4)
library(lmerTest)
library(emmeans)
library(MuMIn)
library(nlme)
```

#### Functions
```{r functions}
# Arc-sine transformation
asinTransform <- function(p) { asin(sqrt(p)) }
# Calculate standard error
sderr <- function(x) {sd(x)/sqrt(length(x))}
#+++++++++++++++++++++++++
# Function to calculate the mean and the standard deviation
# for each group
#+++++++++++++++++++++++++
# data : a data frame
# varname : the name of a column containing the variable
#to be summariezed
# groupnames : vector of column names to be used as
# grouping variables
data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sderr(x[[col]]), na.rm=TRUE)
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
  return(data_sum)
}

normality.plots <- function(x) {
  par(mfrow=c(2,2))
  hist(residuals(x), main = "Histogram", xlab = "Values")
  boxplot(residuals(x), main = "Boxplot", ylab = "Values")
  qqnorm(residuals(x))
  qqline(residuals(x))
  plot(density(residuals(x)), main = "Kernel Density Estimate")
}
```

#### Load the data
Load the alphadiv.csv files created in the RAPID_alpha_statsA.Rmd script. First starting with all species
```{r load}
# load alpha diversity datasets for with and without aquarickettsia
alphadiv <- read.csv("../data/alphadiv_8663.csv")
alphadiv2 <- read.csv("../data/alphadiv_823.csv")
alphadiv2 <- alphadiv2[,1:5]
colnames(alphadiv2) <- c("X","observeda", "chao1a","simpsona","shannona")
alphadiv <- cbind(alphadiv, alphadiv2)
alphadiv <- alphadiv[,-17]
head(alphadiv)

#write.csv(alphadiv, file = "../data/alphadiv_all.csv")
#alphadiv <- read.csv(file = "../data/alphadiv_all.csv")

alphadiv <- alphadiv[which(alphadiv$geno.num != 20),]
```


### 2 Data Summary statistics

```{r}
# Data summaries with Aquarickettsia
# Observed
data_summary(alphadiv, varname = "observed", groupnames = c("bleach"))
data_summary(alphadiv, varname = "observed", groupnames = c("type"))
data_summary(alphadiv, varname = "observed", groupnames = c("bleach.type"))
# Chao1
data_summary(alphadiv, varname = "chao1", groupnames = c("bleach"))
data_summary(alphadiv, varname = "chao1", groupnames = c("type"))
data_summary(alphadiv, varname = "chao1", groupnames = c("bleach.type"))
# Simpson
data_summary(alphadiv, varname = "simpson", groupnames = c("bleach"))
data_summary(alphadiv, varname = "simpson", groupnames = c("type"))
data_summary(alphadiv, varname = "simpson", groupnames = c("bleach.type"))
# Shannon
data_summary(alphadiv, varname = "shannon", groupnames = c("bleach"))
data_summary(alphadiv, varname = "shannon", groupnames = c("type"))
data_summary(alphadiv, varname = "shannon", groupnames = c("bleach.type"))
```

#### Data summaries without Aquarickettsia
```{r}
# Observed
data_summary(alphadiv, varname = "observeda", groupnames = c("bleach"))
data_summary(alphadiv, varname = "observeda", groupnames = c("type"))
data_summary(alphadiv, varname = "observeda", groupnames = c("bleach.type"))
# Chao1
data_summary(alphadiv, varname = "chao1a", groupnames = c("bleach"))
data_summary(alphadiv, varname = "chao1a", groupnames = c("type"))
data_summary(alphadiv, varname = "chao1a", groupnames = c("bleach.type"))
# Simpson
data_summary(alphadiv, varname = "simpsona", groupnames = c("bleach"))
data_summary(alphadiv, varname = "simpsona", groupnames = c("type"))
data_summary(alphadiv, varname = "simpsona", groupnames = c("bleach.type"))
# Shannon
data_summary(alphadiv, varname = "shannona", groupnames = c("bleach"))
data_summary(alphadiv, varname = "shannona", groupnames = c("type"))
data_summary(alphadiv, varname = "shannona", groupnames = c("bleach.type"))
```
### 3 Non Parametric Tests for significance
Using Kruskal Wallis tests

#### Observed ASVs
```{r}
#Kruskal Wallis tests
kruskal.test(observed ~ bleach, data = alphadiv)
kruskal.test(observed ~ type, data = alphadiv)
kruskal.test(observed ~ bleach.type, data = alphadiv)
pairwise.wilcox.test(alphadiv$observed, alphadiv$bleach.type, p.adjust.method = 'fdr')
```

#### Chao1
```{r}
#Kruskal Wallis tests
kruskal.test(chao1 ~ bleach, data = alphadiv)
kruskal.test(chao1 ~ type, data = alphadiv)
kruskal.test(chao1 ~ bleach.type, data = alphadiv)
pairwise.wilcox.test(alphadiv$chao1, alphadiv$bleach.type, p.adjust.method = 'fdr')
```

#### Simpson's Index
```{r}
kruskal.test(simpson ~ bleach, data = alphadiv)
kruskal.test(simpson ~ type, data = alphadiv)
kruskal.test(simpson ~ bleach.type, data = alphadiv)
pairwise.wilcox.test(alphadiv$simpson, alphadiv$bleach.type, p.adjust.method = 'fdr')
```

#### Shannon's Index
```{r}
kruskal.test(shannon ~ bleach, data = alphadiv)
kruskal.test(shannon ~ type, data = alphadiv)
kruskal.test(shannon ~ bleach.type, data = alphadiv)
pairwise.wilcox.test(alphadiv$shannon, alphadiv$bleach.type, p.adjust.method = 'fdr')
```

### 3 Non Parametric Tests for significance without Aquarickettsia
Using Kruskal Wallis tests

#### Observed ASVs
```{r}
#Kruskal Wallis tests
kruskal.test(observeda ~ bleach, data = alphadiv)
kruskal.test(observeda ~ type, data = alphadiv)
kruskal.test(observeda ~ bleach.type, data = alphadiv)
pairwise.wilcox.test(alphadiv$observeda, alphadiv$bleach.type, p.adjust.method = 'fdr')
```

#### Chao1
```{r}
#Kruskal Wallis tests
kruskal.test(chao1a ~ bleach, data = alphadiv)
kruskal.test(chao1a ~ type, data = alphadiv)
kruskal.test(chao1a ~ bleach.type, data = alphadiv)
pairwise.wilcox.test(alphadiv$chao1a, alphadiv$bleach.type, p.adjust.method = 'fdr')
```

#### Simpson's Index
```{r}
kruskal.test(simpsona ~ bleach, data = alphadiv)
kruskal.test(simpsona ~ type, data = alphadiv)
kruskal.test(simpsona ~ bleach.type, data = alphadiv)
pairwise.wilcox.test(alphadiv$simpsona, alphadiv$bleach.type, p.adjust.method = 'fdr')
```

#### Shannon's Index
```{r}
kruskal.test(shannona ~ bleach, data = alphadiv)
kruskal.test(shannona ~ type, data = alphadiv)
kruskal.test(shannona ~ bleach.type, data = alphadiv)
pairwise.wilcox.test(alphadiv$shannona, alphadiv$bleach.type, p.adjust.method = 'fdr')
```
<!-- ### Tests for differences within resistant genotypes -->
<!-- ```{r} -->
<!-- alphadiv_res <- subset(alphadiv, alphadiv$type == "resistant") -->
<!-- kruskal.test(observed ~ bleach, data = alphadiv_res) -->
<!-- kruskal.test(simpson ~ bleach, data = alphadiv_res) -->
<!-- kruskal.test(shannon ~ bleach, data = alphadiv_res) -->
<!-- kruskal.test(faithPD ~ bleach, data = alphadiv_res) -->
<!-- ``` -->

<!-- ### 4 Parametric tests for significance -->
<!-- Looking for normality and homogeneity of group dispersions -->
<!-- ```{r} -->
<!-- # observed ASVs -->
<!-- histogram(alphadiv$observed) -->
<!-- qqnorm(alphadiv$observed) -->
<!-- shapiro.test(alphadiv$observed) -->
<!-- # sqrt transform makes the data normal -->
<!-- histogram(sqrt(alphadiv$observed)) -->
<!-- qqnorm(sqrt(alphadiv$observed)) -->
<!-- shapiro.test(sqrt(alphadiv$observed)) -->
<!-- # test for homogeneity of group dispersions -->
<!-- bartlett.test(sqrt(observed) ~ bleach, data = alphadiv) -->
<!-- bartlett.test(sqrt(observed) ~ geno, data = alphadiv) -->
<!-- ## not significant indicates that the variances in each of the groups are the same -->

<!-- # proceeding with anova tests -->
<!-- summary(aov(sqrt(observed) ~ geno*bleach, alphadiv)) # results disagree with  -->
<!-- summary(aov(sqrt(observed) ~ type, alphadiv)) -->
<!-- #the kruskal wallis test -->
<!-- TukeyHSD(aov(sqrt(observed) ~ geno, alphadiv)) # this gets wonky because too many comparisons -->
<!-- ``` -->

<!-- ```{r} -->
<!-- # Shannon -->
<!-- histogram(alphadiv$shannon) -->
<!-- qqnorm(alphadiv$shannon) -->
<!-- shapiro.test(alphadiv$shannon) -->
<!-- # sqrt transform makes the data normal -->
<!-- histogram(log(alphadiv$shannon)) -->
<!-- qqnorm(log(alphadiv$shannon)) -->
<!-- shapiro.test(log(alphadiv$shannon)) -->
<!-- ## can't seem to find a transformation that makes this normal -->
<!-- ## going to stick with kruskal wallis -->
<!-- ``` -->