---
title: "Muller_beta_stats"
author: "Becca Maher"
date: "November 08, 2019"
output: html_document
---

## 1 Set up
```{r, include=FALSE}
## clear workspace------------------------
rm(list=ls())

# load libraries
library('phyloseq'); packageVersion('phyloseq')
library('vegan'); packageVersion('vegan')
library("usedist")
library("ggplot2")
library("microbiome")

# set working directory-------------------
#setwd("~/Box Sync/RAPID-analysis/")

## functions----------------------
pairwise.adonis.dm <- function(x,factors,stratum=NULL,p.adjust.m="bonferroni",perm=999){
  
  library(vegan)
  
  if(class(x) != "dist") stop("x must be a dissimilarity matrix (dist object)")
  
  co = as.matrix(combn(unique(factors),2))
  
  pairs = c()
  
  F.Model =c()
  
  R2 = c()
  
  p.value = c()
  
  
  
  for(elem in 1:ncol(co)){
    
    sub_inds <- factors %in% c(as.character(co[1,elem]),as.character(co[2,elem]))
    
    resp <- as.matrix(x)[sub_inds,sub_inds]
    
    ad = adonis(as.dist(resp) ~
                  
                  factors[sub_inds], strata=stratum[sub_inds], permutations=perm);
    
    pairs = c(pairs,paste(co[1,elem],'vs',co[2,elem]));
    
    F.Model =c(F.Model,ad$aov.tab[1,4]);
    
    R2 = c(R2,ad$aov.tab[1,5]);
    
    p.value = c(p.value,ad$aov.tab[1,6])
    
  }
  
  p.adjusted = p.adjust(p.value,method=p.adjust.m)
  
  pairw.res = data.frame(pairs,F.Model,R2,p.value,p.adjusted)
  
  return(pairw.res)
  }
```

```{r}
## Data Analysis----------------------------

# First import data from bioinformatic processing
load(file = "../data/ps_rar8692.RData")
ps
sample_sums(ps)
```

## Analysis

### Calculate distance matrices

```{r}
# Calculate all four distance matrices: weighted unifrac, unweighted unifrac, bray curtis, binary jaccard
ps_wu <- phyloseq::distance(ps, method = "wunifrac")
ps_un <- phyloseq::distance(ps, method = "unifrac")
ps_bc <- phyloseq::distance(ps, method = "bray")
ps_bj <- distance(ps, method = "jaccard", binary =TRUE)
```
### PERMANOVAs
```{r}
# PERMANOVA's with Adonis -  Permutational Multivariate Analasis of Variance
# Make a data frame from the sample_data
sampledf <- data.frame(sample_data(ps))
```
#### geno
```{r}
# Adonis individual tests
adonis(ps_bc ~ bleach*geno, data = sampledf)
adonis(ps_bc ~ bleach*type, data = sampledf)
adonis(ps_bj ~ type*geno*bleach, data = sampledf)
adonis(ps_wu ~ type*geno*bleach, data = sampledf)
adonis(ps_un ~ type*geno*bleach, data = sampledf)
# all significant
## pairwise tests are all not significant (padj) probably too many comparisons
## maybe I should use the unadjusted pval?
pairwise.adonis.dm(ps_bc, sample_data(ps)$geno, p.adjust.m = "fdr")
#pairwise.adonis.dm(ps_bj, sample_data(ps)$geno, p.adjust.m = "fdr")
#pairwise.adonis.dm(ps_wu, sample_data(ps)$geno, p.adjust.m = "fdr")
#pairwise.adonis.dm(ps_un, sample_data(ps)$geno, p.adjust.m = "fdr")
```

### PERMDISPs
#### bleach
```{r}
anova(betadisper(ps_bc, sampledf$bleach, bias.adjust = TRUE))
anova(betadisper(ps_bj, sampledf$bleach, bias.adjust = TRUE))
anova(betadisper(ps_wu, sampledf$bleach, bias.adjust = TRUE))
anova(betadisper(ps_un, sampledf$bleach, bias.adjust = TRUE))
# All non-significant
```
#### genotype
```{r}
anova(betadisper(ps_bc, sampledf$geno, bias.adjust = TRUE))
anova(betadisper(ps_bj, sampledf$geno, bias.adjust = TRUE))
anova(betadisper(ps_wu, sampledf$geno, bias.adjust = TRUE))
anova(betadisper(ps_un, sampledf$geno, bias.adjust = TRUE))
```
#### type
```{r}
anova(betadisper(ps_bc, sampledf$type, bias.adjust = TRUE))
anova(betadisper(ps_bj, sampledf$type, bias.adjust = TRUE))
anova(betadisper(ps_wu, sampledf$type, bias.adjust = TRUE))
anova(betadisper(ps_un, sampledf$type, bias.adjust = TRUE))
```
#### Interaction
```{r}
anova(betadisper(ps_bc, sampledf$bleach.geno, bias.adjust = TRUE))
anova(betadisper(ps_bj, sampledf$bleach.geno, bias.adjust = TRUE))
anova(betadisper(ps_un, sampledf$bleach.geno, bias.adjust = TRUE))
anova(betadisper(ps_wu, sampledf$bleach.geno, bias.adjust = TRUE))
# pairwise tests
#p.adjust(permutest(betadisper(ps_un, sampledf$geno.trt, 
#                              bias.adjust = TRUE), 
#                   pairwise=TRUE)$pairwise$permuted, method = 'fdr')
#p.adjust(permutest(betadisper(ps_wu, sampledf$geno.trt, 
#                              bias.adjust = TRUE), 
#                   pairwise=TRUE)$pairwise$permuted, method = 'fdr')

```

Plotting
```{r}
myCol <- c('#e6194b', '#3cb44b', '#ffe119', 
           '#4363d8', '#f58231', '#911eb4', '#46f0f0', 
           '#f032e6', '#bcf60c', '#008080',
           '#9a6324', '#800000', '#808000', 
          '#000075', '#808080', '#000000')


nmds_bc <- metaMDSiter(ps_bc, k=2, trymax = 1000, maxit = 1000, autotransform=FALSE)
#nmds_wu <- metaMDSiter(ps_wu, k=2, trymax = 1000, maxit = 1000, autotransform=FALSE)
#save(nmds_bc, file = "../data/nmds_bc.RData")




a <- plot_ordination(ps, nmds_bc, color = "geno.num", shape = "bleach") + 
  geom_point(size = 2) + 
  theme_classic() +
  theme(legend.position = "left") +
  labs(color = "Genotype") +
  scale_colour_manual(values = myCol)
#a + stat_ellipse(type = "norm", mapping = aes(fill = bleach))
```

## Analysis of log transformed data
Optional transformations to the data
```{r}
# Log-transform OTU table
ps <- transform(ps, transform = "log10")
```

### Calculate distance matrices

```{r}
# Calculate all four distance matrices: weighted unifrac, unweighted unifrac, bray curtis, binary jaccard
ps_wu <- phyloseq::distance(ps, method = "wunifrac")
ps_un <- phyloseq::distance(ps, method = "unifrac")
ps_bc <- phyloseq::distance(ps, method = "bray")
ps_bj <- distance(ps, method = "jaccard", binary =TRUE)
```
### PERMANOVAs
```{r}
# PERMANOVA's with Adonis -  Permutational Multivariate Analasis of Variance
# Make a data frame from the sample_data
sampledf <- data.frame(sample_data(ps))
psdf <- as.data.frame(otu_table(ps))
```
#### geno
```{r}
# Adonis individual tests
adonis(ps_bc ~ bleach*geno, data = sampledf)
adonis(ps_bc ~ bleach*type, data = sampledf)
adonis(ps_bj ~ type*geno*bleach, data = sampledf)
adonis(ps_wu ~ type*geno*bleach, data = sampledf)
adonis(ps_un ~ type*geno*bleach, data = sampledf)
# all significant
## pairwise tests are all not significant (padj) probably too many comparisons
## maybe I should use the unadjusted pval?
pairwise.adonis.dm(ps_bc, sample_data(ps)$geno, p.adjust.m = "fdr")
#pairwise.adonis.dm(ps_bj, sample_data(ps)$geno, p.adjust.m = "fdr")
#pairwise.adonis.dm(ps_wu, sample_data(ps)$geno, p.adjust.m = "fdr")
#pairwise.adonis.dm(ps_un, sample_data(ps)$geno, p.adjust.m = "fdr")
```

### PERMDISPs
#### bleach
```{r}
anova(betadisper(ps_bc, sampledf$bleach, bias.adjust = TRUE))
anova(betadisper(ps_bj, sampledf$bleach, bias.adjust = TRUE))
anova(betadisper(ps_wu, sampledf$bleach, bias.adjust = TRUE))
anova(betadisper(ps_un, sampledf$bleach, bias.adjust = TRUE))
# All non-significant
```
#### genotype
```{r}
anova(betadisper(ps_bc, sampledf$geno, bias.adjust = TRUE))
anova(betadisper(ps_bj, sampledf$geno, bias.adjust = TRUE))
anova(betadisper(ps_wu, sampledf$geno, bias.adjust = TRUE))
anova(betadisper(ps_un, sampledf$geno, bias.adjust = TRUE))
```
#### Interaction
```{r}
anova(betadisper(ps_bc, sampledf$bleach.geno, bias.adjust = TRUE))
anova(betadisper(ps_bj, sampledf$bleach.geno, bias.adjust = TRUE))
anova(betadisper(ps_un, sampledf$bleach.geno, bias.adjust = TRUE))
anova(betadisper(ps_wu, sampledf$bleach.geno, bias.adjust = TRUE))
# pairwise tests
#p.adjust(permutest(betadisper(ps_un, sampledf$geno.trt, 
#                              bias.adjust = TRUE), 
#                   pairwise=TRUE)$pairwise$permuted, method = 'fdr')
#p.adjust(permutest(betadisper(ps_wu, sampledf$geno.trt, 
#                              bias.adjust = TRUE), 
#                   pairwise=TRUE)$pairwise$permuted, method = 'fdr')

```

Plotting
```{r}
myCol <- c('#e6194b', '#3cb44b', '#ffe119', 
           '#4363d8', '#f58231', '#911eb4', '#46f0f0', 
           '#f032e6', '#bcf60c', '#008080',
           '#9a6324', '#800000', '#808000', 
          '#000075', '#808080', '#000000')


nmds_bc <- metaMDSiter(ps_bc, k=2, trymax = 1000, maxit = 1000, autotransform=FALSE)
#nmds_wu <- metaMDSiter(ps_wu, k=2, trymax = 1000, maxit = 1000, autotransform=FALSE)
#save(nmds_bc, file = "../data/nmds_bc.RData")




a <- plot_ordination(ps, nmds_bc, color = "geno.num", shape = "bleach") + 
  geom_point(size = 2) + 
  theme_classic() +
  theme(legend.position = "left") +
  labs(color = "Genotype") +
  scale_colour_manual(values = myCol)