---
title: "Muller_beta_stats"
author: "Becca Maher"
date: "November 08, 2019"
output: html_document
---

## 1 Set up
```{r, include=FALSE}
## clear workspace------------------------
rm(list=ls())

# load libraries
library('phyloseq'); packageVersion('phyloseq')
library('vegan'); packageVersion('vegan')
library("usedist")

# set working directory-------------------
#setwd("~/Box Sync/RAPID-analysis/")

## functions----------------------
pairwise.adonis.dm <- function(x,factors,stratum=NULL,p.adjust.m="bonferroni",perm=999){
  
  library(vegan)
  
  if(class(x) != "dist") stop("x must be a dissimilarity matrix (dist object)")
  
  co = as.matrix(combn(unique(factors),2))
  
  pairs = c()
  
  F.Model =c()
  
  R2 = c()
  
  p.value = c()
  
  
  
  for(elem in 1:ncol(co)){
    
    sub_inds <- factors %in% c(as.character(co[1,elem]),as.character(co[2,elem]))
    
    resp <- as.matrix(x)[sub_inds,sub_inds]
    
    ad = adonis(as.dist(resp) ~
                  
                  factors[sub_inds], strata=stratum[sub_inds], permutations=perm);
    
    pairs = c(pairs,paste(co[1,elem],'vs',co[2,elem]));
    
    F.Model =c(F.Model,ad$aov.tab[1,4]);
    
    R2 = c(R2,ad$aov.tab[1,5]);
    
    p.value = c(p.value,ad$aov.tab[1,6])
    
  }
  
  p.adjusted = p.adjust(p.value,method=p.adjust.m)
  
  pairw.res = data.frame(pairs,F.Model,R2,p.value,p.adjusted)
  
  return(pairw.res)
  }
```

```{r}
## Data Analysis----------------------------

# First import data from bioinformatic processing
load(file = "./data/muller_bleaching_silva_species_assign.RData")
map = import_qiime_sample_data("./data/map.txt")
sample_names(map)
sample_data(ps) <- map
ps
```

## 2 Analysis of log transformed data

Optional transformations to the data
```{r}
# Log-transform OTU table
otus_log <- as(otu_table(ps), "matrix")
otus_log <- decostand(otus_log, method = "log")
OTU_log <- otu_table(otus_log, taxa_are_rows = TRUE)
otu_table(ps) <- OTU_log
```

### Calculate distance matrices

```{r}
# Calculate all four distance matrices: weighted unifrac, unweighted unifrac, bray curtis, binary jaccard
ps_wu <- phyloseq::distance(ps, method = "wunifrac")
ps_un <- phyloseq::distance(ps, method = "unifrac")
ps_bc <- phyloseq::distance(ps, method = "bray")
ps_bj <- distance(ps, method = "jaccard", binary =TRUE)
```
### PERMANOVAs
```{r}
# PERMANOVA's with Adonis -  Permutational Multivariate Analasis of Variance
# Make a data frame from the sample_data
sampledf <- data.frame(sample_data(ps))
psdf <- as.data.frame(otu_table(ps))
```
#### geno
```{r}
# Adonis individual tests
adonis(ps_bc ~ geno, data = sampledf)
adonis(ps_bj ~ geno, data = sampledf)
adonis(ps_wu ~ geno, data = sampledf)
adonis(ps_un ~ geno, data = sampledf)
# all significant
pairwise.adonis.dm(ps_bc, sample_data(ps)$geno, p.adjust.m = "fdr")
pairwise.adonis.dm(ps_bj, sample_data(ps)$geno, p.adjust.m = "fdr")
pairwise.adonis.dm(ps_wu, sample_data(ps)$geno, p.adjust.m = "fdr")
pairwise.adonis.dm(ps_un, sample_data(ps)$geno, p.adjust.m = "fdr")
```
#### bleach
```{r}
adonis(ps_bc ~ bleach, data = sampledf)
adonis(ps_bj ~ bleach, data = sampledf)
adonis(ps_wu ~ bleach, data = sampledf)
adonis(ps_un ~ bleach, data = sampledf)
# all non-significant
```
#### Species
```{r}
# Adonis won't run on samples that have an NA for species
# So I subset the dist to get rid of the samples without a species ID
samples_wna <- c("July2016.228.2", "July2016.251.2", "July2016.251.1")
samples <- sample_names(ps)
samples_nona <- setdiff(samples, samples_wna)
ps_bc_sub <- dist_subset(ps_bc, samples_nona)
ps_bj_sub <- dist_subset(ps_bj, samples_nona)
ps_wu_sub <- dist_subset(ps_wu, samples_nona)
ps_un_sub <- dist_subset(ps_un, samples_nona)

adonis(ps_bc_sub ~ species, data = sampledf)
adonis(ps_bj_sub ~ species, data = sampledf)
adonis(ps_wu_sub ~ species, data = sampledf)
adonis(ps_un_sub ~ species, data = sampledf)

sampledf_na <- subset(sampledf, sample.id != "July2016.228.2" & sample.id != "July2016.251.2" & sample.id != "July2016.251.1")

# All significant
pairwise.adonis.dm(ps_bc_sub, sampledf_na$species, p.adjust.m = "fdr")
pairwise.adonis.dm(ps_bj_sub, sampledf_na$species, p.adjust.m = "fdr")
pairwise.adonis.dm(ps_wu_sub, sampledf_na$species, p.adjust.m = "fdr")
pairwise.adonis.dm(ps_un_sub, sampledf_na$species, p.adjust.m = "fdr")
```
#### Nutrient
bleach was control or nutrient (urea or nitrate)
```{r}
adonis(ps_bc ~ nutrient, data = sampledf)
adonis(ps_bj ~ nutrient, data = sampledf)
adonis(ps_wu ~ nutrient, data = sampledf)
adonis(ps_un ~ nutrient, data = sampledf)
# No significant results
```
#### Interaction
```{r}
# Adonis test for between group diversity, with full formula
adonis(ps_bc_sub ~ geno*species*bleach, data = sampledf) # sig: geno, species, geno:species
adonis(ps_bj_sub ~ geno*species*bleach, data = sampledf) # sig: geno, species, geno:species
adonis(ps_wu_sub ~ geno*species*bleach, data = sampledf) # sig: geno, species
adonis(ps_un_sub ~ geno*species*bleach, data = sampledf) # sig: geno, species, geno:species
```

### PERMDISPs
#### Species
```{r}
# PERMDISP with betadisper - Multivariate Homogeneity of group dispersions
anova(betadisper(ps_bc, sampledf$species, bias.adjust = TRUE)) # non sig
anova(betadisper(ps_bj, sampledf$species, bias.adjust = TRUE)) # non sig
anova(betadisper(ps_wu, sampledf$species, bias.adjust = TRUE)) # significant
anova(betadisper(ps_un, sampledf$species, bias.adjust = TRUE)) # non sig
# pairwise test
p.adjust(permutest(betadisper(ps_wu, sampledf$species, 
                              bias.adjust = TRUE), 
                   pairwise=TRUE)$pairwise$permuted, method = 'fdr')
```
#### bleach
```{r}
anova(betadisper(ps_bc, sampledf$bleach, bias.adjust = TRUE))
anova(betadisper(ps_bj, sampledf$bleach, bias.adjust = TRUE))
anova(betadisper(ps_wu, sampledf$bleach, bias.adjust = TRUE))
anova(betadisper(ps_un, sampledf$bleach, bias.adjust = TRUE))
# All non-significant
```
#### geno
```{r}
anova(betadisper(ps_bc, sampledf$geno, bias.adjust = TRUE)) # non sig
anova(betadisper(ps_bj, sampledf$geno, bias.adjust = TRUE)) # significant
anova(betadisper(ps_wu, sampledf$geno, bias.adjust = TRUE)) # significant
anova(betadisper(ps_un, sampledf$geno, bias.adjust = TRUE)) # significant
# pairwise tests
p.adjust(permutest(betadisper(ps_bj, sampledf$geno, 
                              bias.adjust = TRUE), 
                   pairwise=TRUE)$pairwise$permuted, method = 'fdr')
p.adjust(permutest(betadisper(ps_wu, sampledf$geno, 
                              bias.adjust = TRUE), 
                   pairwise=TRUE)$pairwise$permuted, method = 'fdr')
p.adjust(permutest(betadisper(ps_un, sampledf$geno, 
                              bias.adjust = TRUE), 
                   pairwise=TRUE)$pairwise$permuted, method = 'fdr')

```
#### Interaction
```{r}
anova(betadisper(ps_bc, sampledf$geno.trt, bias.adjust = TRUE))
anova(betadisper(ps_bj, sampledf$geno.trt, bias.adjust = TRUE))
anova(betadisper(ps_un, sampledf$geno.trt, bias.adjust = TRUE)) # significant
anova(betadisper(ps_wu, sampledf$geno.trt, bias.adjust = TRUE)) # significant
# pairwise tests
p.adjust(permutest(betadisper(ps_un, sampledf$geno.trt, 
                              bias.adjust = TRUE), 
                   pairwise=TRUE)$pairwise$permuted, method = 'fdr')
p.adjust(permutest(betadisper(ps_wu, sampledf$geno.trt, 
                              bias.adjust = TRUE), 
                   pairwise=TRUE)$pairwise$permuted, method = 'fdr')

```

## 3 Analysis of log transformed Acropora data
```{r}
# subset by species
ps <- subset_samples(ps, species =="ACR")
#ps <- subset_samples(ps, species =="POC")
#ps <- subset_samples(ps, geno == "T2")
```

Calculate distance matrices
```{r}
# Calculate all four distance matrices: weighted unifrac, unweighted unifrac, bray curtis, binary jaccard
ps_wu <- phyloseq::distance(ps, method = "wunifrac")
ps_un <- phyloseq::distance(ps, method = "unifrac")
ps_bc <- phyloseq::distance(ps, method = "bray")
ps_bj <- distance(ps, method = "jaccard", binary =TRUE)
```

### PERMANOVAs
```{r cars}
# PERMANOVA's with Adonis -  Permutational Multivariate Analasis of Variance
# Make a data frame from the sample_data
sampledf <- data.frame(sample_data(ps))
```
#### geno
```{r}
# Adonis individual tests
adonis(ps_bc ~ geno, data = sampledf)
adonis(ps_bj ~ geno, data = sampledf)
adonis(ps_wu ~ geno, data = sampledf)
adonis(ps_un ~ geno, data = sampledf)
# all significant
# pairwise tests
pairwise.adonis.dm(ps_bc, sample_data(ps)$geno, p.adjust.m = "fdr")
pairwise.adonis.dm(ps_bj, sample_data(ps)$geno, p.adjust.m = "fdr")
pairwise.adonis.dm(ps_wu, sample_data(ps)$geno, p.adjust.m = "fdr")
pairwise.adonis.dm(ps_un, sample_data(ps)$geno, p.adjust.m = "fdr")
```
#### bleach
```{r}
adonis(ps_bc ~ bleach, data = sampledf)
adonis(ps_bj ~ bleach, data = sampledf)
adonis(ps_wu ~ bleach, data = sampledf)
adonis(ps_un ~ bleach, data = sampledf)
# all non significant
```
#### Interaction
```{r}
adonis(ps_bc ~ geno*bleach, data = sampledf) # geno
adonis(ps_bj ~ geno*bleach, data = sampledf) # geno
adonis(ps_wu ~ geno*bleach, data = sampledf) # geno
adonis(ps_un ~ geno*bleach, data = sampledf) # geno
```
### PERMDISPs
#### bleach
```{r}
# PERMDISP with betadisper - Multivariate Homogeneity of group dispersions
anova(betadisper(ps_bc, sampledf$bleach, bias.adjust = TRUE)) # significant
anova(betadisper(ps_bj, sampledf$bleach, bias.adjust = TRUE))
anova(betadisper(ps_wu, sampledf$bleach, bias.adjust = TRUE))
anova(betadisper(ps_un, sampledf$bleach, bias.adjust = TRUE))
# pairwise
p.adjust(permutest(betadisper(ps_bc, sampledf$bleach, 
                              bias.adjust = TRUE), 
                   pairwise=TRUE)$pairwise$permuted, method = 'fdr')
```
#### geno
```{r}
anova(betadisper(ps_bc, sampledf$geno, bias.adjust = TRUE))
anova(betadisper(ps_bj, sampledf$geno, bias.adjust = TRUE)) # sig
anova(betadisper(ps_wu, sampledf$geno, bias.adjust = TRUE)) # sig
anova(betadisper(ps_un, sampledf$geno, bias.adjust = TRUE))
# pairwise
p.adjust(permutest(betadisper(ps_bj, sampledf$geno, 
                              bias.adjust = TRUE), 
                   pairwise=TRUE)$pairwise$permuted, method = 'fdr')
p.adjust(permutest(betadisper(ps_wu, sampledf$geno, 
                              bias.adjust = TRUE), 
                   pairwise=TRUE)$pairwise$permuted, method = 'fdr')
```
Plotting
```{r}
nmds_bc <- metaMDSiter(ps_bc, k=2, trymax = 1000, maxit = 1000, autotransform=FALSE)
nmds_bc <- metaMDSiter(ps_bc, k=2, trymax = 1000, maxit = 1000, autotransform=FALSE)

a <- plot_ordination(ps, nmds_bc, color = "bleach") + geom_point(size = 2) + theme_classic() +
  theme(legend.position = "left") + scale_colour_colorblind()
a

```{r}